Survey of Cache Oblivious B-trees

### 2011-12-14 paper.tex
Tweaking figure placement
Making the paper more 'professional' (*rollseyes*) at Leilani's request. Also fixing the header on one section and fixing a couple of the citations.
added a description section that is not a one-liner to variable length strings for consistency

@@

\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Salman Ahmad\\
\affaddr{MIT CSAIL}
\email{saahmad@mit.edu}
% 2nd. author
\alignauthor
Leilani Battle\\
\affaddr{MIT CSAIL}
\email{leibatt@mit.edu}
% 3rd. author
\alignauthor
Stephen Tu\\
\affaddr{MIT CSAIL}
\email{stephent@mit.edu}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
\date{7 December 2011}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

@@

\begin{center}
	\includegraphics[width=0.8\columnwidth]{figures/veb.pdf}
\end{center}

@@

@@

\begin{center}
	\includegraphics[width=0.8\columnwidth]{figures/buffers.pdf}
\end{center}

\caption{An illustration of how the shuttle tree determines the size of a node's linked list. The 
node $n2$ will have a linked list of size 2 because it is the leaf of two of the recursive subtrees. 
On the other hand $n1$ will only have a linked list of size 1.}
\label{fig:buffers}
\end{figure}

@@

\section{Variable Length Strings}
\begin{figure}

\begin{center}
	\includegraphics[width=0.8\columnwidth]{figures/COSB-treegeneralstructurediagram.pdf}
\end{center}

\caption{An illustration of the dynamic COSB-tree structure}
\label{fig:string_structure}
\end{figure}
\subsection{Motivation}

@@

\subsubsection{General COSB-tree Structure}

The dynamic COSB-tree consists of three layers: a centroid tree, a PMA 
of hash data for the keys (known as the 
\emph{hashdata} layer), and a PMA of compressed values for the original key
 (known as the 
\emph{keydata} layer). Each leaf of the centroid tree contains a pointer into
the hashdata array. The centroid tree contains only hash values, and is designed such that
each leaf points to the respective \emph{neighborhood} of the key being
searched for. After traversing the centroid tree and following the pointer into
the hashdata array layer, we are within $O(\log{}N)$ keys of $k$. Then a
sequential scan is performed in the hashdata array layer to find $k$. Each
entry in the hashdata array contains only the hash value of the corresponding key,
and a pointer to the compressed key value in the keydata layer. The keydata layer
is consulted to resolve hash collisions in the hashdata layer, and to obtain the original value of the key.

@@

\subsubsection{Centroid Trees} %todo:cite dictionary matching problem--sources in cosb-tree paper
\begin{figure}

\begin{center}
	\includegraphics[width=0.8\columnwidth]{figures/centroidtreeexample.pdf}
\end{center}

\caption{A centroid tree example from \cite{BenderFaKu06}. Dotted lines
represent centroid-rooted up tries, where the parent is not a prefix of the
child. Solid lines represent centroid-rooted down tries, where the parent is a
prefix of the child.}
\label{fig:centroid}
\end{figure}

@@

@@

\begin{center}
	\includegraphics[width=0.8\columnwidth]{figures/centroidrelayoutexample.pdf}
\end{center}

@@

@@

@@

\begin{figure*}[h!t]
\begin{center}
    \begin{tabular}{ |   l   |  l     |     l     |     }
    \hline
	Data Structure & Search & Insert / Delete  \\ \hline \hline
	 \textit{Original CO B-tree}  & $O(\log_{B}{N})$ & $O(\log_{B}{N} + \frac{\log^2{N}}{B})$  \\ \hline
	\textit{String CO B-tree}  & $O(1+\log_{B}N+||k||/B)$ &   $O(1+\log_{B}N+\log^{2}N||k||/B)$  \\ \hline
	\textit{Streaming Shuttle Tree}  & $O(\log_{B}{N})$ &  $O(log_{B+1} N / B^{\Theta(1/(loglogB)^2)} + (log^2 N) / B)$  \\ \hline
	\textit{Streaming COLA}  & $O(\log N)$ & $O((\log N) / B)$  \\ \hline
	\textit{Concurrent Exponential CO B-tree} &   $O(\log_B{N} + \log_{\alpha}{\log{B}})$  &   $O(\log_B{N} + \log{\alpha}{\log{B}})$  \\ \hline
	\textit{Concurrent PMA CO B-tree}  & $O(\log_{B}{N} + ({\log{N}})/{B} + 1)$ &   $O(\log_{B}{N} + (\log^2{N})/{B} + 1)$  \\ \hline
    \end{tabular}
    \caption{A summary of the surveyed CO B-trees. The bounds are in terms of the number of memory transfers. The bounds for \textit{Concurrent Exponential CO B-tree} are in expectation only.}
    \label{fig:table}
\end{center}
\end{figure*}

@@

@@

\section{Acknowledgments}
We sincerely thank Dr. David Karger and all of the TAs for 6.854 
for their guidance, help, and support. May the Flow (Max-Flow, that is) be with you all.

>>>>

\addtolength{\topmargin}{-.07in}
\addtolength{\textheight}{.07in}

%\addtolength{\oddsidemargin}{-0.08in}
%\addtolength{\evensidemargin}{-0.08in}
%\addtolength{\textwidth}{0.08in}

@@

\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Salman Ahmad\\
\affaddr{MIT Media Lab}
\email{saahmad@mit.edu}
% 2nd. author
\alignauthor
Leilani Battle\\
\affaddr{MIT CSAIL}
\email{leibatt@mit.edu}
% 3rd. author
\alignauthor
Stephen Tu\\
\affaddr{MIT CSAIL}
\email{stephent@mit.edu}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
\date{7 December 2011}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

@@

\begin{center}
	\includegraphics[width=0.75\columnwidth]{figures/veb.pdf}
\end{center}

@@

\vspace{-0.1in}

@@

\begin{figure}
\begin{center}
	\includegraphics[width=0.8\columnwidth]{figures/buffers.pdf}
\end{center}
\caption{An illustration of how the shuttle tree determines the size of a node's linked list. The 
node $n2$ will have a linked list of size 2 because it is the leaf of two of the recursive subtrees. 
On the other hand $n1$ will only have a linked list of size 1.}
\label{fig:buffers}
\end{figure}

@@

\section{Variable Length Strings}

@@

\subsection{Description}
The following subsections describe the structure of the COSB-tree in more detail,
as well as summarize the modifications Bender et al. make to front compression
for COSB-trees.
\subsubsection{General COSB-tree Structure}

\begin{figure}
\begin{center}
	\includegraphics[width=\columnwidth]{figures/COSB-treegeneralstructurediagram.pdf}
\end{center}
\caption{An illustration of the dynamic COSB-tree structure}
\label{fig:string_structure}
\end{figure}

The dynamic COSB-tree consists of three layers: a centroid tree, a PMA 
of hash data for the keys (known as the 
\emph{hashdata} layer), and a PMA of compressed values for the original key
 (known as the 
\emph{keydata} layer). Each leaf of the centroid tree contains a pointer into
the hashdata array. The centroid tree contains only hash values, and is designed such that
each leaf points to the respective \emph{neighborhood} of the key being
searched for. After traversing the centroid tree and following the pointer into
the hashdata array layer, we are within $O(\log{}N)$ keys of $k$. Then a
sequential scan is performed in the hashdata array layer to find $k$. Each
entry in the hashdata array contains only the hash value of the corresponding key,
and a pointer to the compressed key value in the keydata layer. The keydata layer
is consulted to resolve hash collisions in the hashdata layer, and to obtain the original value of the key.
The structure is described in Figure~\ref{fig:string_structure}.

@@

@@

\subsubsection{Centroid Trees} %todo:cite dictionary matching problem--sources in cosb-tree paper
\begin{figure}
\begin{center}
	\includegraphics[width=0.8\columnwidth]{figures/centroidtreeexample.pdf}
\end{center}
\caption{A centroid tree example from \cite{BenderFaKu06}. Dotted lines
represent centroid-rooted up tries, where the parent is not a prefix of the
child. Solid lines represent centroid-rooted down tries, where the parent is a
prefix of the child.}
\label{fig:centroid}
\end{figure}

@@

\begin{center}
	\includegraphics[width=0.7\columnwidth]{figures/centroidrelayoutexample.pdf}
\end{center}

@@

\begin{figure*}[t]
\begin{center}
    \begin{tabular}{ |   l   |  l     |     l     |     }
    \hline
	Data Structure & Search & Insert / Delete  \\ \hline \hline
	 \textit{Original CO B-tree}  & $O(\log_{B}{N})$ & $O(\log_{B}{N} + \frac{\log^2{N}}{B})$  \\ \hline
	\textit{String CO B-tree}  & $O(1+\log_{B}N+||k||/B)$ &   $O(1+\log_{B}N+\log^{2}N||k||/B)$  \\ \hline
	\textit{Streaming Shuttle Tree}  & $O(\log_{B}{N})$ &  $O(log_{B+1} N / B^{\Theta(1/(loglogB)^2)} + (log^2 N) / B)$  \\ \hline
	\textit{Streaming COLA}  & $O(\log N)$ & $O((\log N) / B)$  \\ \hline
	\textit{Concurrent Exponential CO B-tree} &   $O(\log_B{N} + \log_{\alpha}{\log{B}})$  &   $O(\log_B{N} + \log{\alpha}{\log{B}})$  \\ \hline
	\textit{Concurrent PMA CO B-tree}  & $O(\log_{B}{N} + ({\log{N}})/{B} + 1)$ &   $O(\log_{B}{N} + (\log^2{N})/{B} + 1)$  \\ \hline
    \end{tabular}
    \caption{A summary of the surveyed CO B-trees. The bounds are in terms of the number of memory transfers. The bounds for \textit{Concurrent Exponential CO B-tree} are in expectation only.}
    \label{fig:table}
\end{center}
\end{figure*}

@@

\vfill

@@

\vfill

@@

\newpage

@@

\vspace{0.1in}

\section{Acknowledgments}

We sincerely thank Dr. David Karger and all of the 6.854 TAs
for their guidance, help, and support.

\vspace{0.1in}

%%%% May the Flow (Max-Flow, that is) be with you all.
